From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-19 Pi+/paper/finite.tex
Add outline of braid diagrams

To do the proof, we are using a presentation based on the adjacent
transpositions, i.e. where the primitive operations are adjacent swaps. There
are three laws that these operations have to satisfy - it is easiest to
visualize them on diagrams.
\begin{itemize}
\item Swapping the same two elements two times in a row is the same as doing nothing:
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, s_1}};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {cancel} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={1, 1}};
    \end{tikzpicture}
\item When swapping two pairs of elements on a positions far away, it does not
matter in which order it is done (We can slide the wires).
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ 1, s_1}};
    \end{tikzpicture}
    \hspace{0cm}
    \dots
    \hspace{0cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, 1 }};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {swap} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, 1}};
    \end{tikzpicture}
    \hspace{0cm}
    \dots
    \hspace{0cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ 1, s_1}};
    \end{tikzpicture}

\item seen on a diagram (braid)
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 3,
        thick]
        {braid={ s_2, s_1, s_2}};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {braid} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 3,
        thick]
        {braid={ s_1, s_2, s_1}};
    \end{tikzpicture}

\end{itemize}

Diagram of long-braid:
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 2,
    thick]
    {braid={1, 1, 1, 1, s_1, 1}};
\end{tikzpicture}
\hspace{0cm}
\dots
\hspace{0cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 5,
    thick]
    {braid={ s_4, s_3, s_2, s_1, 1, s_4}};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
    \draw [-stealth](0,0) -- node[anchor=south] {long-braid} (2,0);
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 2,
    thick]
    {braid={1, 1, 1, 1, 1, s_1}};
\end{tikzpicture}
\hspace{0cm}
\dots
\hspace{0cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 5,
    thick]
    {braid={ s_3, s_4, s_3, s_2, s_1, 1}};
\end{tikzpicture}
\hspace{1cm}

>>>>

We use a presentation based on adjacent transpositions, which is called the
Coxeter presentation. The primitive operations are adjacent swaps. Further,
there are three laws that these operations have to satisfy -- it is easiest to
visualise them as braid diagrams.

Swapping the same two elements two times in a row is the same as doing nothing:

\[
  \begin{tabular}{m{0.3\linewidth}m{0.1\linewidth}m{0.3\linewidth}}
    \begin{center}
      \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd,
          number of strands = 2,
          thick]
        {braid={ s_1, s_1}};
      \end{tikzpicture}
    \end{center}
     &
    \(\xlongrightarrow[]{\cancel}\)
     &
    \begin{center}
      \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd,
          number of strands = 2,
          thick]
        {braid={1, 1}};
      \end{tikzpicture}
    \end{center}
  \end{tabular}
\]

When swapping two pairs of elements in positions far away, it does not matter in
which order the swapping happens, that is, we can slide the wires.

\[
  \begin{tabular}{m{0.3\linewidth}m{0.1\linewidth}m{0.3\linewidth}}
    \begin{center}
      \begin{tabular}{m{0.3\linewidth}m{0.1\linewidth}m{0.3\linewidth}}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={ 1, s_1}};
        \end{tikzpicture}
         &
        \(\cdots\)
         &
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={ s_1, 1 }};
        \end{tikzpicture}
      \end{tabular}
    \end{center}
     &
    \(\xlongrightarrow[]{\swap}\)
     &
    \begin{center}
      \begin{tabular}{m{0.3\linewidth}m{0.1\linewidth}m{0.3\linewidth}}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={ s_1, 1}};
        \end{tikzpicture}
         &
        \(\cdots\)
         &
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={ 1, s_1}};
        \end{tikzpicture}
      \end{tabular}
    \end{center}
  \end{tabular}
\]

\[
  \begin{tabular}{m{0.3\linewidth}m{0.1\linewidth}m{0.3\linewidth}}
    \begin{center}
      \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd,
          number of strands = 3,
          thick]
        {braid={ s_2, s_1, s_2}};
      \end{tikzpicture}
    \end{center}
     &
    \(\xlongrightarrow[]{\braid}\)
     &
    \begin{center}
      \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd,
          number of strands = 3,
          thick]
        {braid={ s_1, s_2, s_1}};
      \end{tikzpicture}
    \end{center}
  \end{tabular}
\]

Diagram of long-braid:

\[
  \begin{tabular}{m{0.45\linewidth}m{0.1\linewidth}m{0.45\linewidth}}
    \begin{tabular}{m{0.15\linewidth}m{0.05\linewidth}m{0.1\linewidth}}
      \begin{center}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={1, 1, 1, 1, s_1, 1}};
        \end{tikzpicture}
      \end{center}
       &
      \(\cdots\)
       &
      \begin{center}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 5,
            thick]
          {braid={ s_4, s_3, s_2, s_1, 1, s_4}};
        \end{tikzpicture}
      \end{center}
    \end{tabular}
     &
    \(\xlongrightarrow[]{\longbraid}\)
     &
    \begin{tabular}{m{0.15\linewidth}m{0.05\linewidth}m{0.1\linewidth}}
      \begin{center}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 2,
            thick]
          {braid={1, 1, 1, 1, 1, s_1}};
        \end{tikzpicture}
      \end{center}
       &
      \(\cdots\)
       &
      \begin{center}
        \begin{tikzpicture}
          \pic[local bounding box=my braid,braid/.cd,
            number of strands = 5,
            thick]
          {braid={ s_3, s_4, s_3, s_2, s_1, 1}};
        \end{tikzpicture}
      \end{center}
    \end{tabular}
  \end{tabular}
\]
